### 2018-03-02 submission/manuscript.Rmd
final re-read edits

The adenoma models as a whole performed poorly in classifying individuals with and without adenomas. This outcome is not inconsistent with what has been published previously [@hale_shifts_2017; @Sze2017]. However, the modeling results are at odds with results obtained in Baxter, *et al.* [@baxter_microbiota-based_2016]. There are some major differences between the models generated in this meta-analysis and what was used in Baxter, *et al*. First, the prevoius report's models investigated the classification of lesions (individuals with adenoma or carcinoma) and not adenoma alone. The Baxter, *et al.* models also contained Fecal Immunoglobulin Test data while our meta-analysis models only contained 16S rRNA gene sequencing data. Although being able to classify individuals with adenomas is important, the classification of advanced adenomas is a more clinically meaningful diagnostic tool (i.e. those that are at high risk of progressing to a carcinoma). It is possible that we might have been able to detect differences in the bacterial community if advanced adenomas were separated from adenomas but that data was not available for the majority of studies analyzed. Additionally, the initial changes to the bacterial community could be focal to where the initial adenoma develops and would not be easily assessed with a fecal sample.

Although stool represents an easy and less invasive way to assess risk, it is not clear how well this sample reflects adenoma- and carcinoma- tissue associated microbial communities. The colon tissue-based studies did not provide a clearer understanding of how the microbiota may be associated with tumors. Generally, the full OTU-based models of unmatched and matched colon tissue samples were concordant with stool samples showing that GI resident microbes were the most prevalent in the top 10 most important variables across study [Figure S3]. *Fusobacterium* was not consistently identified across studies and this could be due to both a small number of studies and a small sample size within these studies. The majority of the colon tissue-based results were also study specific with many of the top 10 taxa being present only in a single study. Additionally, the presence of genera associated with contamination [@Salter_contamination_2014], within the top 10 most important variables for the genera and OTU models is worrying (e.g. *Novosphingobium*, *Acidobacteria Gp2*, *Sphingomonas*, etc.). The low bacterial biomass of tissue samples coupled with potential contamination and small sample sizes could explain why these results seem to be more sporadic than the stool results.

>>>>

The adenoma models as a whole performed poorly in classifying individuals with and without adenomas. This outcome is not inconsistent with what has been published previously [@hale_shifts_2017; @Sze2017]. However, the modeling results are at odds with results obtained in Baxter, *et al.* [@baxter_microbiota-based_2016]. There are some major differences between the models generated in this meta-analysis and what was used in Baxter, *et al*. First, the prevoius report's models investigated the classification of lesions (individuals with adenoma or carcinoma) and not adenoma alone. The Baxter, *et al.* models also contained Fecal Immunoglobulin Test data while our meta-analysis models only contained 16S rRNA gene sequencing data. Although being able to classify individuals with adenomas is important, the classification of advanced adenomas is a more clinically meaningful diagnostic tool (i.e. those that are at highest risk of progressing to a carcinoma). It is possible that we might have been able to detect differences in the bacterial community if advanced adenomas were separated from adenomas but that data was not available for the majority of studies analyzed. Additionally, the initial changes to the bacterial community could be focal to where the initial adenoma develops and would not be easily assessed with a fecal sample.

Although stool represents an easy and less invasive way to assess risk, it is not clear how well this sample reflects adenoma- and carcinoma- tissue associated microbial communities. The colon tissue-based studies did not provide a clearer understanding of how the microbiota may be associated with tumors. Generally, the full OTU-based models of unmatched and matched colon tissue samples were concordant with stool samples showing that GI resident microbes were the most prevalent in the top 10 most important variables across study [Figure S3]. *Fusobacterium* was not consistently identified across studies and this could be due to both a small number of studies and a small sample size within these studies. The majority of the colon tissue-based results were also study specific with many of the top 10 taxa being present only in a single study. Additionally, the presence of genera associated with contamination, within the top 10 most important variables for the genera and OTU models is worrying (e.g. *Novosphingobium*, *Acidobacteria Gp2*, *Sphingomonas*, etc. [@Salter_contamination_2014]). The low bacterial biomass of tissue samples coupled with potential contamination and small sample sizes could explain why these results seem to be more sporadic than the stool results.
